Open-source landscape for 3D CSEM modelling

### 2020-06-19 manuscript/Manuscript.tex

\itodo{
We will have to decide on a Journal. Current suggestions:\\
- GMD: Geoscientific Model Development\\
- SIG: Surveys in Geophysics\\
- SE: Solid Earth\\
- GJI: Geophysical Journal International\\
- GP: Geophysical Prospecting\\
- GEO: Geophysics\\
- CAG: Computers and Geosciences\\
I think \emph{Geoscientific Model Development} and \emph{Surveys in Geophysics}
were the favourite ones when we discussed it ones. I personally would be very
interested in GMD. What do you think? But also interesting would be the GJI,
because the MT comparison was published there, so it would be in the same
place. I just had a very bad experience with GP, so I am not inclined on that
at all.
}

\mycom[RR]{GMD looks very interesting, my only concern is that it might not be
in the scope of EM modelers, but we can probably reach modeleres from other
communities as well.}
\mycom[OCR]{I also find GMD very interesting, in fact it was my first option to publish the first petgem version (2 years ago). The answer was: "manuscript outside of journal's scope". In this sense I agree with Raphael. My vote is for SIG}

@@

\imycom[OCR]{In general, I believe that the content is adequate and the ideas are clear and in accordance with the objective that we have initially set. However, I want to highlight the task of being careful when justifying the absence of the SimPEG results for the marlim model (I have read it in slack, although it may not be definitive). I believe that this may or may not be a weak point depending on how we explain it. I know that perhaps it is not serious, I only make this comment so that the speech justifies the facts. }

@@

\mycom[OCR]{I have rearranged some references according to the year of publication (here and in the rest of the paper)}
The continuous improvement of computing power and the CSEM boom in the early
2000's in the hydrocarbon industry led to a wealth of developed numerical
solutions and according publications. The most common applied methods to solve
Maxwell's equation are the IE method \citep{GJI.74.Raiche, RS.02.Hursan,
GEO.06.Zhdanov, GP.10.Tehrani, CAG.16.Kruglyakov, MGS.17.Kruglyakov} and
different variations of the differential equation (DE) method, such as finite
differences (FD) \citep{IEEE.66.Yee, GEO.93.Wang, RSC.94.Mackie, RS.94.Druskin,
GEO.09.Streich, CAG.13.Sommer}, finite elements (FE) \citep{GEO.04.Commer,
GJI.11.Schwarzbach,GEO.12.daSilva, GJI.13.Grayver, GJI.13.Puzyrev, SEG.16.Zhang},
and finite volumes (FV) \citep{EM.90.Madsen, PIER.01.Clemens, GP.06.Mulder, ECP.07.Haber, GEO.14.Jahandari}. There are also many different types of
discretisation, where the most common ones are regular grids (cartesian,
rectilinear), mostly using a Yee grid \citep{IEEE.66.Yee} or a Lebedev grid
\citep{CMMP.64.Lebedev}, but also unstructured tetrahedral grids
\citep{SEG.16.Zhang, CAG.17.Cai}, hexagonal meshes \citep{CAG.14.Cai}, or
OcTree meshes\citep{ECP.07.Haber}. Very well written overviews about the
different approaches to 3D EM modelling are given by \cite{SG.05.Avdeev} and
\cite{SG.10.Borner}.

@@

The meaning of the term open-source evolved rapidly in the last two decades.
Today, open-source generally not only means that a code is available (with a
proper license), but also that the code is hosted online, is under version
control, has the possibility to file issues and make pull requests to fix bugs.
These are important requirements for reproducible research «\emph{to provide
the means by which the reader can verify the validity of the results and make
use of them in further research}» \citep{GEO.17.Broggini}. Well maintained
codes often have continuous integration that includes testing of the code base
and online hosted documentation, often including examples and tutorials. As
such, the term has shifted from purely open-source codes to developments in the
open with increasingly building a global community. Related projects within the
realm of geophysical exploration are e.g., \texttt{pyGIMLi}
\citep{CAG.17.Rucker} or \texttt{Fatiando} \citep{JOSS.18.Uieda}.
\imycom[DW]{Others coming to your mind?\imycom[RR]{no thoughts about geophysics,
  but there are many projects for related topics such as gmt, gis, mesh-gen}
  \imycom[OCR]{No idea from my side}}

@@

\tabl[btp]{codecomparison}{Comparison of the four codes under consideration.
  Note that \emg3d is a solver on its own, while the other codes implement
  third-package solvers such as \texttt{PETSc} \citep{Preprint.Abhyankar},
  \texttt{MUMPS} \citep{SIAM.01.Amestoy}, or \texttt{PARDISO}
  \citep{FGCS.04.Schenk}.
}{
  \centering
  \footnotesize
\begin{tabularx}{\linewidth}{lXXXX}
  \toprule
  %
  & \custem & \emg3d & \petgem & \simpeg  \\
  \midrule
  %
  Home & \href{https://custem.rtfd.io}{custem.rtfd.io}
       & \href{https://empymod.github.io}{empymod.github.io}
       & \href{http://petgem.bsc.es}{petgem.bsc.es}
       & \href{https://docs.simpeg.xyz}{simpeg.xyz} \\
  License & GPL-3.0 & Apache-2.0 & GPL-3.0 & MIT \\
  Installation & \texttt{conda}
               & \texttt{pip}; \texttt{conda}
               & \texttt{pip}; \texttt{conda}
               & \texttt{pip}; \texttt{conda} \\
  Comp. Dom. & frequency \& time & frequency & frequency & frequency \& time \\
  Method & FE & FV & FE & FV \\
  Mesh & tetrahedral & rectilinear & tetrahedral & recti- \& curvilinear \\
  BC & PMC; PEC & PEC & PEC & PEC; PMC \\
  Solver & \texttt{MUMPS} & \emg3d & \texttt{PETSc}; \texttt{MUMPS} &
           \texttt{PARDISO}; \texttt{MUMPS} \\
  %
  \bottomrule
\end{tabularx}}%
%

%
\tabl[btp]{machines}{List of computer and operating system (hardware and
software) on which the different codes were run, together with the operator.}{
  \centering
  \footnotesize
  \begin{tabularx}{\linewidth}{lXl}
  \toprule
  %
  Code & Computer and Operating System & Operator \\
  \midrule
  %
  \custem & PowerEdge R940 server; 144 Xeon Gold 6154 CPU @2.666 GHz; ~3 TB
            DDR4 RAM; Ubuntu 18.04
          & Raphael Rochlitz \\
  \emg3d  & Laptop; i7-6600U CPU@2.6 GHz x4; 16 GB of memory, Ubuntu 18.04
          & Dieter Werthmüller \\
  \petgem & Marenostrum4. Intel Xeon Platinum from Skylake generation; 2
            sockets Intel Xeon Platinum 8160 CPU with 24 cores each @2.10GHz
            for a total of 48 cores per node; 386 Gb DDR4 RAM per node; SuSE
            Linux Enterprise
          & Octavio Castillo-Reyes \\
  \simpeg & ???
          & Lindsey Heagy \\
  %
  \bottomrule
\end{tabularx}}%
%

@@

\petgem permits to locate the source and receivers anywhere in the computational domain (e.g., sediments, seafloor, sea, ground), allowing to analyze the physical environment of the electric responses and how parameters impact them (e.g., frequency, conductivity, dependence on mesh setup, basis order, solver type, among others). Furthermore, \petgem implements a semi-adaptive mesh strategy ($hp$ mesh refinement) based on physical parameters and on polynomial order to satisfy quality criteria chosen by the user. Nonetheless, only Horizontal electric dipole (HED) has been implemented.

@@

The Marlim oil field is a giant reservoir in a turbidite sandstone horizon in
the north-eastern part of the Campos Basin, offshore Brazil, which was
discovered in 1985. \cite{BJG.17.Carvalho} created from seismic data and well
log data a realistic resistivity model with vertical transverse isotropy (VTI),
Marlim R3D (MR3D), which they released under an open-source license.
\cite{GEO.19.Correa} computed CSEM data for six frequencies from 0.125\,Hz to
1.25\,Hz, released under an open-source license. To compute the data they used
a state-of-the-art code from the industry \citep{GEO.07.Maao}. It is therefore
an ideal case to verify our open-source codes against, as it is a complex,
realistic model and the data were computed by an industry-proofed
code. Additionally, that code is a time-domain code, whereas the four codes
under consideration here compute the results all in the frequency-domain.
\imycom[RR]{I would delete this sentence, not that important and a bit
redundant.\imycom[DW]{Did you refer to the last sentence or the one before? I
think the last sentence contains quite important info. .\imycom[OCR]{If possible, I would like to uphold this sentence. I agree with dieter, it contains relevant information.}}}

The detailed reservoir model consists of 1022 by 371 by 1229 cells\mycom[OCR]{I prefer this notation $1022 \times 371 \times 1229$. Nothing special, just format}, totalling
to almost 466 million cells, where each cell has dimensions of 25 by 75 by
5\,m. For the computation the model is upscaled to 515 by 563 by 310 cells,
totalling to almost 90 million cells, where each cell has dimensions of 100 by
100 by 20\,m.

@@

\imycom[DW]{I would like to capture as many as possible of the (I) differences,
  (II) difficulties, (III) advantages, and (IV) disadvantages, that we
  encountered in this exercise. This includes differences from direct to
  iterative solver and differences from FE to FD codes etc. Please just jot
  down things that come to your mind.
  \imycom[OCR]{ I think an important point is the user experience (modeler). I know this is an intangible element but perhaps we could add a line about it. It is my opinion}}

@@

The work of D.W. was conducted within the Gitaro.JIM project funded through
MarTERA, a EU Horizon 2020 research and innovation programme (No 728053);
\href{https://www.martera.eu}{martera.eu}. The development of \custem as part of
the DESMEX project was funded by the Germany Ministry for Education and
Research (BMBF) in the framework of the research and development program
Fona-r4 under grant 033R130D. The work of O.C-R. has received funding from the \textit{European Union\textquotesingle s Horizon 2020 programme} under the \textit{Marie Sklodowska-Curie} grant agreement N$^\circ$ 777778. Further, the development of \petgem has received funding from the \textit{European Union\textquotesingle s Horizon 2020 programme}, grant agreement N$^\circ$ 828947, and from the Mexican Department of Energy, CONACYT-SENER Hidrocarburos grant agreement N$^\circ$ B-S-69926. Furthermore, O.C-R. has been 65\% cofinanced by the European Regional Development Fund (ERDF) through the Interreg V-A Spain-France-Andorra program (POCTEFA2014-2020). POCTEFA aims to reinforce the economic and social integration of the French-Spanish-Andorran border. Its support is focused on developing economic, social and environmental cross-border activities through joint strategies favouring sustainable territorial development.\mycom[OCR]{My apologies for these long acks, but they are the projects in which I am participating and the BSC makes a lot of emphasis on complying verbatim with the thansk sentences (too rigorous in my opinion)}

>>>>

\itodo{
We will have to decide on a Journal. Current suggestions:\\
- GMD: Geoscientific Model Development\\
- SIG: Surveys in Geophysics\\
- SE: Solid Earth\\
- GJI: Geophysical Journal International\\
- GP: Geophysical Prospecting\\
- GEO: Geophysics\\
- CAG: Computers and Geosciences\\
\imycom[DW]{
I think \emph{Geoscientific Model Development} and \emph{Surveys in Geophysics}
were the favourite ones when we discussed it ones. I personally would be very
interested in GMD. What do you think? But also interesting would be the GJI,
because the MT comparison was published there, so it would be in the same
place. I just had a very bad experience with GP, so I am not inclined on that
at all.}
\imycom[RR]{GMD looks very interesting, my only concern is that it might not be
in the scope of EM modelers, but we can probably reach modeleres from other
communities as well.}
\imycom[OCR]{I also find GMD very interesting, in fact it was my first option to
publish the first petgem version (2 years ago). The answer was: "manuscript
outside of journal's scope". In this sense I agree with Raphael. My vote is for
SIG}
}

@@

\imycom[OCR]{In general, I believe that the content is adequate and the ideas
are clear and in accordance with the objective that we have initially set.
However, I want to highlight the task of being careful when justifying the
absence of the SimPEG results for the Marlim model (I have read it in Slack,
although it may not be definitive). I believe that this may or may not be a
weak point depending on how we explain it. I know that perhaps it is not
serious, I only make this comment so that the speech justifies the facts. }

@@

The continuous improvement of computing power and the CSEM boom in the early
2000's in the hydrocarbon industry led to a wealth of developed numerical
solutions and according publications. The most common applied methods to solve
Maxwell's equation are the IE method \citep{GJI.74.Raiche, RS.02.Hursan,
GEO.06.Zhdanov, GP.10.Tehrani, CAG.16.Kruglyakov, MGS.17.Kruglyakov} and
different variations of the differential equation (DE) method, such as finite
differences (FD) \citep{IEEE.66.Yee, GEO.93.Wang, RSC.94.Mackie, RS.94.Druskin,
GEO.09.Streich, CAG.13.Sommer}, finite elements (FE) \citep{GEO.04.Commer,
GJI.11.Schwarzbach,GEO.12.daSilva, GJI.13.Grayver, GJI.13.Puzyrev,
SEG.16.Zhang}, and finite volumes (FV) \citep{EM.90.Madsen, PIER.01.Clemens,
GP.06.Mulder, ECP.07.Haber, GEO.14.Jahandari}. There are also many different
types of discretisation, where the most common ones are regular grids
(cartesian, rectilinear), mostly using a Yee grid \citep{IEEE.66.Yee} or a
Lebedev grid \citep{CMMP.64.Lebedev}, but also unstructured tetrahedral grids
\citep{SEG.16.Zhang, CAG.17.Cai}, hexagonal meshes \citep{CAG.14.Cai}, or
OcTree meshes\citep{ECP.07.Haber}. Very well written overviews about the
different approaches to 3D EM modelling are given by \cite{SG.05.Avdeev} and
\cite{SG.10.Borner}.

@@

The meaning of the term open-source evolved rapidly in the last two decades.
Today, open-source generally not only means that a code is available (with a
proper license), but also that the code is hosted online, is under version
control, has the possibility to file issues and make pull requests to fix bugs.
These are important requirements for reproducible research «\emph{to provide
the means by which the reader can verify the validity of the results and make
use of them in further research}» \citep{GEO.17.Broggini}. Well maintained
codes often have continuous integration that includes testing of the code base
and online hosted documentation, often including examples and tutorials. As
such, the term has shifted from purely open-source codes to developments in the
open with increasingly building a global community. Related projects within the
realm of geophysical exploration are e.g., \texttt{pyGIMLi}
\citep{CAG.17.Rucker} or \texttt{Fatiando} \citep{JOSS.18.Uieda}.
\imycom[DW]{Others coming to your mind? An interesting looking one is
  \href{https://github.com/fancompute/ceviche}{github.com/fancompute/ceviche},
  but they are at very high frequencies on nm-distances.
\imycom[RR]{no thoughts about geophysics, but there are many projects for
  related topics such as gmt, gis, mesh-gen}
\imycom[OCR]{No idea from my side}
}

@@

\tabl[btp]{codecomparison}{Comparison of the four codes under consideration.
  Note that \emg3d is a solver on its own, while the other codes implement
  third-package solvers such as \texttt{PETSc} \citep{Preprint.Abhyankar},
  \texttt{MUMPS} \citep{SIAM.01.Amestoy}, or \texttt{PARDISO}
  \citep{FGCS.04.Schenk}.
}{
  \centering
  \footnotesize
\begin{tabularx}{\linewidth}{lXXXX}
  \toprule
  %
  & \custem & \emg3d & \petgem & \simpeg  \\
  \midrule
  %
  Home & \href{https://custem.rtfd.io}{custem.rtfd.io}
       & \href{https://empymod.github.io}{empymod.github.io}
       & \href{http://petgem.bsc.es}{petgem.bsc.es}
       & \href{https://docs.simpeg.xyz}{simpeg.xyz} \\
  License & GPL-3.0 & Apache-2.0 & GPL-3.0 & MIT \\
  Installation & \texttt{conda}
               & \texttt{pip}; \texttt{conda}
               & \texttt{pip}
               & \texttt{pip}; \texttt{conda} \\
  Comp. Dom. & frequency \& time & frequency & frequency & frequency \& time \\
  Method & FE & FV & FE & FV \\
  Mesh & tetrahedral & rectilinear & tetrahedral & recti- \& curvilinear \\
  BC & PMC; PEC & PEC & PEC & PEC; PMC \\
  Solver & \texttt{MUMPS} & \emg3d & \texttt{PETSc}; \texttt{MUMPS} &
           \texttt{PARDISO}; \texttt{MUMPS} \\
  %
  \bottomrule
\end{tabularx}}%
%

%
\tabl[btp]{machines}{List of computer and operating system (hardware and
software) on which the different codes were run, together with the operator.}{
  \centering
  \footnotesize
  \begin{tabularx}{\linewidth}{lXl}
  \toprule
  %
  Code & Computer and Operating System & Operator \\
  \midrule
  %
  \custem & \textbf{PowerEdge R940 server}; 144 Xeon Gold 6154 CPU @2.666 GHz;
            ~3 TB DDR4 RAM; Ubuntu 18.04
          & Raphael Rochlitz \\
  \emg3d  & \textbf{Laptop}; i7-6600U CPU@2.6 GHz x4; 16 GB of memory,
            Ubuntu 18.04 & Dieter Werthmüller \\
  \petgem & \textbf{Marenostrum4 (supercomputer)}. Intel Xeon Platinum from
            Skylake generation; 2 sockets Intel Xeon Platinum 8160 CPU with 24
            cores each @2.10GHz for a total of 48 cores per node; 386 Gb DDR4
            RAM per node; SuSE Linux Enterprise
          & Octavio Castillo-Reyes \\
  \simpeg & ???
          & Lindsey Heagy \\
  %
  \bottomrule
\end{tabularx}}%
%

@@

\petgem permits to locate the source and receivers anywhere in the
computational domain\mycom[DW]{I think this applies to all four codes} (e.g.,
sediments, seafloor, sea, ground), allowing to analyze the physical environment
of the electric responses and how parameters impact them (e.g., frequency,
conductivity, dependence on mesh setup, basis order, solver type, among
others). Furthermore, \petgem implements a semi-adaptive mesh strategy ($hp$
mesh refinement) based on physical parameters and on polynomial order to
satisfy quality criteria chosen by the user. Nonetheless, only Horizontal
Electric Dipole (HED) has been implemented.

@@

The Marlim oil field is a giant reservoir in a turbidite sandstone horizon in
the north-eastern part of the Campos Basin, offshore Brazil, which was
discovered in 1985. \cite{BJG.17.Carvalho} created from seismic data and well
log data a realistic resistivity model with vertical transverse isotropy (VTI),
Marlim R3D (MR3D), which they released under an open-source license.
\cite{GEO.19.Correa} computed CSEM data for six frequencies from 0.125\,Hz to
1.25\,Hz, released under an open-source license. To compute the data they used
a state-of-the-art code from the industry \citep{GEO.07.Maao}. It is therefore
an ideal case to verify our open-source codes against, as it is a complex,
realistic model and the data were computed by an industry-proofed code.
Additionally, that code is a time-domain code, whereas the four codes under
consideration here compute the results all in the frequency-domain.
\imycom[RR]{I would delete this sentence, not that important and a bit
  redundant.
\imycom[DW]{Did you refer to the last sentence or the one before? I think the
  last sentence contains quite important info.}
\imycom[OCR]{If possible, I would like to uphold this sentence. I agree with
  Dieter, it contains relevant information.}
}

The detailed reservoir model consists of $1022 \times 371 \times 1229$ cells,
totalling to almost 466 million cells, where each cell has dimensions of 25 by
75 by 5\,m. For the computation the model is upscaled to 515 by 563 by 310
cells, totalling to almost 90 million cells, where each cell has dimensions of
100 by 100 by 20\,m.

@@

\imycom[DW]{I would like to capture as many as possible of the (I) differences,
  (II) difficulties, (III) advantages, and (IV) disadvantages, that we
  encountered in this exercise. This includes differences from direct to
  iterative solver and differences from FE to FD codes etc. Please just jot
  down things that come to your mind.
\imycom[OCR]{I think an important point is the user experience (modeler). I
  know this is an intangible element but perhaps we could add a line about it.
  It is my opinion}
}

@@

The work of D.W. was conducted within the Gitaro.JIM project funded through
MarTERA, a EU Horizon 2020 research and innovation programme (No 728053);
\href{https://www.martera.eu}{martera.eu}. The development of \custem as part of
the DESMEX project was funded by the Germany Ministry for Education and
Research (BMBF) in the framework of the research and development program
Fona-r4 under grant 033R130D. The work of O.C-R. has received funding from the
\emph{European Union's Horizon 2020 programme} under the \emph{Marie
Sklodowska-Curie} grant agreement N$^\circ$ 777778. Further, the development of
\petgem has received funding from the \emph{European Union's Horizon 2020
programme}, grant agreement N$^\circ$ 828947, and from the Mexican Department
of Energy, CONACYT-SENER Hidrocarburos grant agreement N$^\circ$ B-S-69926.
Furthermore, O.C-R. has been 65\% cofinanced by the European Regional
Development Fund (ERDF) through the Interreg V-A Spain-France-Andorra program
(POCTEFA2014-2020). POCTEFA aims to reinforce the economic and social
integration of the French-Spanish-Andorran border. Its support is focused on
developing economic, social and environmental cross-border activities through
joint strategies favouring sustainable territorial development.
\imycom[OCR]{My
  apologies for these long acks, but they are the projects in which I am
  participating and the BSC makes a lot of emphasis on complying verbatim with
  the thansk sentences (too rigorous in my opinion)}
